\texttt{nuance}: Efficient detection of planets transiting active stars

### 2024-07-19 latex/paper.tex
add all figures to same folder + snakemake ref + icons for online ressources and source code

\documentclass[modern,linenumbers]{aastex631}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{bm}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{float}
\usepackage{soul}

% commands
\newcommand{\nuancemethod}{\textit{nuance}}
\newcommand{\nuance}{\nuancemethod{}}
\newcommand{\nuancecode}{\textsf{nuance}}

\newcommand{\TODO}{\texttt{todo}}
\newcommand{\review}[1]{#1}
% \newcommand{\review}[1]{#1} 
\newcommand{\set}[1]{\{\,#1\,\}}
\newcommand{\footlink}[1]{\footnote{\url{#1}}}
\newcommand{\wtls}{\texttt{bi-weight+BLS}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\let\subsectionautorefname\sectionautorefname
\let\subsubsectionautorefname\sectionautorefname

% no indent
\setlength\parindent{0pt}

% code style
\lstdefinestyle{mystyle}{
    backgroundcolor=\color{white},
    commentstyle=\color{gray!70},
    keywordstyle=\color{Bittersweet},
    stringstyle=\color{RoyalBlue},
    basicstyle=\fontsize{8.5}{13}\fontfamily{DejaVuSansMono-TLF}\selectfont,
    breakatwhitespace=false,         
    breaklines=true,
    rulecolor=\color{black!15},
    numbers=none,
    numberstyle=\fontsize{7}{11}\fontfamily{DejaVuSansMono-TLF}\selectfont\color{gray!50},
    framerule=0pt,
    breakindent=5pt,
    resetmargins=true,
    numbersep=10pt,
    frame=single,
    aboveskip=1em,
    belowskip=1em,
    xleftmargin=6pt,
    framexleftmargin=4pt
}
\lstset{style=mystyle}

@@

% author info
\author{Lionel J. Garcia}
\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY, USA}
\affiliation{\review{Astrobiology Research Unit, University of Liège, Liège, Belgium}{}}
\author{Daniel Foreman-Mackey}
\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY, USA}
\author{Catriona A. Murray}
\affiliation{Department of Astrophysical and Planetary Sciences, University of Colorado, Boulder, USA}
\author{Suzanne Aigrain}
\affiliation{Sub-department of Astrophysics, Department of Physics, University of Oxford, Oxford, UK}
\author{Dax L. Feliz}
\affiliation{American Museum of Natural History, New York, NY, USA}
\author{Francisco J. Pozuelos}
\affiliation{Instituto de Astrofísica de Andalucía, Granada, Spain}
\affiliation{Astrobiology Research Unit, University of Liège, Liège, Belgium}
\correspondingauthor{Lionel J. Garcia}
\email{lgarcia@flatironinstitute.org}

@@

\begin{abstract}
    The detection of planetary transits in the light curves of active stars, featuring correlated noise in the form of stellar variability, remains a challenge. Depending on the noise characteristics, we show that the traditional technique that consists of detrending a light curve before searching for transits alters their signal-to-noise ratio, and hinders our capability to discover exoplanets transiting rapidly-rotating active stars. We present \nuancemethod{}, an algorithm to search for transits in light curves while simultaneously accounting for the presence of correlated noise, such as stellar variability and instrumental signals. We assess the performance of \nuancemethod{} on simulated light curves as well as on the TESS light curves of 438 rapidly-rotating M dwarfs. For each dataset, we compare our method to 5 commonly-used detrending techniques followed by a search with the Box-Least-Squares algorithm. Overall, we demonstrate that \nuancemethod{} is the most performant method in 93\% of cases, leading to both the highest number of true positives and the lowest number of false positive detections. Although simultaneously searching for transits while modeling correlated noise is expected to be computationally expensive, we make our algorithm tractable and available as the \textsf{JAX}-powered Python package \href{https://github.com/lgrcia/nuance}{\nuancecode{}}, allowing its use on distributed environments and GPU devices. Finally, we explore the prospects offered by the \nuancemethod{} formalism, and its use to advance our knowledge of planetary systems around active stars, both using space-based surveys and sparse ground-based observations.
\end{abstract}

@@

In this paper, we identify regions of light curves morphological parameter space for which a full-fledged transit search is necessary, and we present \nuancemethod{}\footnote{Throughout the paper, \nuancemethod{} written in italics refers to the algorithm, while \nuancecode{} in sans-serif refers to its implementation published as a Python package in its version 0.5.2 (see \autoref{package}).}, a method to search for transit signals while simultaneously modeling correlated noises in a tractable way. In \autoref{issues}, we describe the effect of correlated noise on transit light curves and the effect of its detrending on transit signals detectability. In \autoref{nuance}, we present \nuance{}, and the two main steps on which this method is based: the linear search and the periodic search. In \autoref{results}, we test the performance of \nuance{} on a wide variety of cases, and compare our method to commonly used transit search algorithms. This include transits injected in synthetic datasets, but also in the TESS light curves of 438 rapidly-rotating M dwarfs. Finally, in \autoref{discussion}, we discuss the results and the limitations of \nuance{}, before concluding in \autoref{conclusion}.

@@

        \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_dataset_decomposed.pdf}
        \caption{The top plot shows the simulated flux time series of an active star drawn from the GP model described in \autoref{light_curves_simulations}. This dataset corresponds to an observation of 4 days with an exposure time of 2 minutes. The mean of this signal consists of a periodic transit signal of period $P=0.7$ days, duration $D=1.2$ hours and depth $\Delta=2\%$ (dark gray) plus an instrumental signal (blue). Correlated noise in the form of stellar variability is simulated with an SHO kernel of hyperparameters $\omega = \pi / 6 D$, $Q=45$ and $\sigma = \Delta$. Finally, white noise with a standard deviation of $0.1\%$ is added to the diagonal of the covariance matrix.}

@@

One way to quantify the detectability of a unique transit signal is to compute its signal-to-noise-ratio expressed as 
\begin{equation*}
  SNR= \frac{\Delta}{\sigma}\sqrt{n},
\end{equation*}
where $\Delta$ is the transit depth, $\sigma$ the measurements uncertainty (assuming homoscedasticity), and $n$ the number of points within transit. Although this metric is useful to assess the strength of the transit signal given a certain photometric precision, it does not account for the presence of correlated noise. However, instrumental and other astrophysical signals will necessarily affect the detectability of transits in realistic light curves \citep{Pont2006}. For this reason, the combined differential photometric precision (CDPP; \citealt{Jenkins2010}) metric was developed, and used in the context of the Kepler mission to assess the level of correlated noise in light curves, affecting the detectability of transits with a given duration. 
% Ultimately, this metric was essential to estimate the yield of the survey, and helped to achieve one of the primary goal of the mission: to measure the occurrence rate of Earth-size planets orbiting in the habitable zones of sun-like stars \citep{}.
The CDPP is computed by decomposing the data in the time-frequency domain using wavelets, and measures the significance of a distorted transit signal in the whitened data. For a given transit duration, the CDPP is then a measure of the noise remaining after filtering the light curve in each frequency band, taking into account the presence of nonstationary correlated noise of a given timescale (see \citealt{Jenkins2010} for more details). Hence, we can estimate the significance of the transit signal accounting for the presence of correlated noise as
\begin{equation}\label{eq:snr}
  SNR= \frac{\hat{\Delta}}{CDPP_D},
\end{equation}
where $CDPP_D$ is the CDPP computed for a given transit duration $D$ and $\hat{\Delta}$ is the transit depth after detrending. For simplicity, the CDPP is computed using the method from \cite{Gilliland2011} implemented in the \textsf{lightkurve} Python package \citep{Lightkurve} and $\hat{\Delta}$ is the depth obtained by solving \autoref{eq:linear_model} with the last column of $\bm{X}$ containing a normalized transit model.
\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=0.6\linewidth]{../workflows/plot_issues/figures/issue1.pdf}
        \caption{Illustration of the effect of correlated noise on a single transit SNR. An hour-long transit of depth 0.5\% is generated on top of white noise (standard deviation of 0.15\%) as part of a 24-hours observation with an exposure time of 1 minute (top). Then, in the bottom plot, correlated noise is added, generated using a GP with an SHO kernel of hyperparameters $\omega=60$, $Q=0.5$ and $\sigma=\Delta/4$ . The SNR on the right of each light curve is computed using \autoref{eq:snr}. Models used to simulate these data are provided in \autoref{light_curves_simulations}.}
        \label{fig:issue1}
    \end{centering}
\end{figure}
\autoref{fig:issue1} shows the SNR from \autoref{eq:snr} computed for a unique transit observed in the absence (gray) and presence of correlated noise (red). As illustrated in this figure, the presence of correlated noise strongly decreases the transit signal SNR, which would ultimately limit its detectability.

@@

        \caption{In each plot on the left, correlated noises simulated using a GP with an SHO kernel are added (blue line) to an original light curve containing a transit of depth $\Delta = 0.8\%$, duration $D=0.1$ day, and white noise with a standard deviation of 0.15\% (top left). The hyperparameters of the kernel are fixed to $Q=10$ and $\sigma=0.8\%$, with the pulsation $\omega$ increasing from top to bottom. The resulting stellar variability signals are then reconstructed using (in purple) Tukey's bi-weight filter with an optimal window size of $3\times D$  (see \citealt{wotan}), and (in red) a GP with the same kernel used to simulate the data. In each case the variability is reconstructed, subtracted, and the transit SNR computed using \autoref{eq:snr}.}

@@

        \caption{Transit SNR computed for 10 000 differential light curves with varying morphological characteristics. All light curves span 2.8 days with an exposure time of 2 minutes and contain the same transit signal of  duration $D=1$ hour, depth of 0.5\% and white noise with a standard deviation of 0.1\%. Light curves at the top and right side of the central plot are shown with their corresponding $\tau$ and $\delta$ values, which again correspond to the relative timescale of stellar variability against the transit duration, and the relative amplitude of stellar variability against the transit depth. The two central plots show the transit SNR after detrending the light curves with one of two methods. On the left, light curves are detrended using Tukey's bi-weight filter with an optimal window size of $3\times D$ (see \citealt{wotan}), while on the right, light curves are detrended using a GP with the same kernel used to simulate the data.}

@@

The goal of the linear search is to compute the likelihood $p(\bm{f} \vert T_0 , D, \Delta)$ for a grid of epochs $\set{T_i}_i$ and durations $\set{D_j}_j$. For each pair $(T_i, D_j)$ the transit depth $\Delta_{i, j}$ is linearly solved, which leads to the set of maximum likelihoods
\begin{equation*}
    \set{\ln\mathcal{L}_{i,j}}_{i, j} = \set{\ln p(\bm{f} \vert T_i ,D_j, \Delta_{i, j})}_{i, j}.
\end{equation*}
An example of such a grid of likelihoods is shown in \autoref{fig:linear_search}. To prepare for the next step, uncertainties on the depths $\sigma_{i,j}$ are also computed using \autoref{eq:ls} and stored.
\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_linear_search.pdf}
        \caption{Principle and output of the linear search. The simulated dataset (top) corresponds to the one shown and described in \autoref{fig:app_principle_dataset}. First, a set of durations and depths $\set{T_i, D_j}_{i,j}$ is generated. For each pair of indices $(i,j)$, the likelihood $\ln p(\bm{f} \vert T_i ,D_j, \Delta_{i,j})$ is computed using the parameters from \autoref{eq:ls} and the expression of \autoref{eq:linear_search_ll}. This process yields the grid of log-likelihoods $\ln\mathcal{L}$ (bottom plot), as well as the $\set{\Delta_{i,j}, \sigma_{i,j}}_{i, j}$ transit depths and errors inferred linearly using \autoref{eq:ls}.}
        \label{fig:linear_search}
    \end{centering}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_periodic_1.pdf}\\
    \caption{Periodic search likelihood $\mathcal{P}(P)$ computed for different trial periods $P$. Notice how the maximum value of $\mathcal{P}$ for the alias period $P=0.35$ day (left plot) is lower than for $P=2\times 0.35=0.7$ day, a result of combining the log-likelihoods using \autoref{eq:result} instead of \autoref{eq:attempt}, in favor of individual transits matching a common depth $\Delta$.}
    \label{fig:periodic_search}
\end{figure}

@@

Using \autoref{eq:result}, we can now compute $\ln\mathcal{P}$ for a range of periods and phases, and build a transit search periodogram using \autoref{eq:periodogram}. This has two disadvantages: First, each likelihood $p(\bm{f} \vert T_0, D, \Delta)$ estimated during the linear search is computed using N measurements. Hence, combining transits in the periodic search, through $\Delta_k$, $\sigma_k$ and the product of $K$ likelihoods $\set{\mathcal{L}_k}_k$ (see \autoref{eq:result}), artificially leads to a likelihood involving up to $N\times K$ measurements. For this reason, one has to normalize each likelihood $\ln\mathcal{P}$, by keeping track of the number of points used to compute each of them, which differs from one phase to another. Second, the maximum value of the likelihood $\ln\mathcal{P}$ is relative to a given dataset, so that a more intuitive and absolute metric must be used to relate to the transit signal detection (such as the Signal Detection Efficiency in \citealt{bls}). This motivates a final step to produce an interpretable transit search periodogram $\mathcal{Q}$. 
\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_Q.pdf}
        \caption{For each period $P$, the joint likelihood $\mathcal{P}(P)$ is computed using \autoref{eq:result}, and the value of the maximum likelihood transit SNR retained as $\mathcal{Q}(P)$.}
        \label{fig:periodogram}
    \end{centering}
\end{figure}
For any period P, instead of taking $ \mathcal{Q}(P)$ as the maximum value of $\ln\mathcal{P}$, we compute the maximum likelihood parameters
\begin{equation}\label{eq:phi0}
    (\phi_0 ,D) = \argmax_{\phi_i, D_j} \set{\ln p(\bm{f} \vert P, \phi_i, D_j)}_{i, j}
\end{equation}
and define $\mathcal{Q}(P)$ as being equal to the SNR of the transit of period $P$, epoch $T_0 = \phi_0 \times P$, duration $D$ and depth $\Delta$, i.e.
\begin{equation*}
    \mathcal{Q}(P) = \frac{\Delta}{\sigma},
\end{equation*}
where $\Delta$ and $\sigma$ are obtained using \autoref{eq:result} with the last column of $X$ containing a periodic transit signal of period $P$, epoch $T_0$, duration $D$ and depth $1$. This process and the resulting periodogram $\mathcal{Q}$ are shown in \autoref{fig:periodogram}.\\\\
Hence, periodic transit of period $P$ with the maximum SNR, i.e. maximizing $\mathcal{Q}$, is adopted as the best candidate, basing the confidence in this signal through its SNR. The parameters of this transit are the period $P$, epoch $T_0 = \phi_0 P$, duration $D$ (\autoref{eq:phi0}), and depth $\Delta$ with error $\sigma$ (given by \autoref{eq:result}).\\\\
\subsection{An open-source python package}\label{package}
The methods presented in this paper are made available through the \nuancecode{} open-source Python package hosted at \href{https://github.com/lgrcia/nuance}{https://github.com/lgrcia/nuance}, released on the Python Package Index\footlink{https://pypi.org/project/nuance/} and with documentation and tutorials hosted at \href{https://nuance.readthedocs.io}{https://nuance.readthedocs.io}. All following mentions of \nuancecode{} refer to its \review{version 0.6.0 \citep{zenodo}.}{} \\\\
To instantiate a search, a user can start by creating a \textsf{Nuance} object with
\begin{lstlisting}[language=Python]
from nuance import Nuance

@@

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=\linewidth]{../workflows/control_test_bls/figures/control_test.pdf}
        \caption{Binned statistics of the injection-recovery of 4000 transit signals in a flat light curve with only white noise using \textsf{BLS} and \nuancecode{}. The color scale indicates the recovery of transits in the corresponding (period, depth) parameter space, white for a full recovery and black for no detection.}
        \label{fig:control}
    \end{centering}
\end{figure}

@@

        \caption{Rate of true positives for the 4000 simulated light curves. The color scale represents the fraction of true positives, white if all injected transits are recovered in a given portion of the $(\tau, \delta)$ parameter space, black if none are recovered.}

@@

    \includegraphics[width=\linewidth]{../workflows/tess_injection_recovery/figures/cleaned/1019692.pdf}
    \caption{Trimmed and cleaned single-sector light curve of the target TIC 1019692. The top plot shows how much of the data is truncated and sigma clipped, resulting in a quasi-continuous light curve shown in the bottom plot. On this bottom plot, the black line corresponds to the mean of the GP model (with hyperparameters optimized here on the cleaned light curve).}

@@

        \caption{Rate of true and false positives for all methods as a fraction of detectable transit signals.}

@@

        \caption{Rate of true and false positives of \nuancecode{} compared to other methods, as a function of the relative parameters $\tau$ and $\delta$. The solid blue line corresponds to the maximum of true positives among all methods on the top panel (dominated \review{by the \texttt{harmonics+BLS} method}), and the minimum of false positives on the bottom panel (waned by the \texttt{GP+BLS} method, all other methods being above \nuancecode{}).}

@@

        \caption{Rate of true and false positives binned over the full parameter space $(\tau, \delta)$ for all methods and all targets.}

@@

        \caption{\review{Rate of true and false positives of \nuancecode{} compared to other methods, as a function of the injected planet's radius. The solid blue line corresponds to the maximum of true positives among all other methods than \nuance{} (dominated by the \texttt{harmonics+BLS} method).}{}}

>>>>

\documentclass[modern]{aastex631}

\include{stylez}

@@

% author info
\author[0000-0002-4296-2246]{Lionel J. Garcia}
\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY, USA}
\affiliation{\review{Astrobiology Research Unit, University of Liège, Liège, Belgium}{}}
\author[0000-0002-9328-5652]{Daniel Foreman-Mackey}
\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY, USA}
\author[0000-0001-8504-5862]{Catriona A. Murray}
\affiliation{Department of Astrophysical and Planetary Sciences, University of Colorado, Boulder, USA}
\author[0000-0003-1453-0574]{Suzanne Aigrain}
\affiliation{Sub-department of Astrophysics, Department of Physics, University of Oxford, Oxford, UK}
\author[0000-0002-2457-7889]{Dax L. Feliz}
\affiliation{American Museum of Natural History, New York, NY, USA}
\author[0000-0003-1572-7707]{Francisco J. Pozuelos}
\affiliation{Instituto de Astrofísica de Andalucía, Granada, Spain}
\affiliation{Astrobiology Research Unit, University of Liège, Liège, Belgium}
\correspondingauthor{Lionel J. Garcia}
\email{lgarcia@flatironinstitute.org}

@@

\begin{abstract}
    The detection of planetary transits in the light curves of active stars, featuring correlated noise in the form of stellar variability, remains a challenge. Depending on the noise characteristics, we show that the traditional technique that consists of detrending a light curve before searching for transits alters their signal-to-noise ratio, and hinders our capability to discover exoplanets transiting rapidly-rotating active stars. We present \nuancemethod{}, an algorithm to search for transits in light curves while simultaneously accounting for the presence of correlated noise, such as stellar variability and instrumental signals. We assess the performance of \nuancemethod{} on simulated light curves as well as on the TESS light curves of 438 rapidly-rotating M dwarfs. For each dataset, we compare our method to 5 commonly-used detrending techniques followed by a search with the Box-Least-Squares algorithm. Overall, we demonstrate that \nuancemethod{} is the most performant method in 93\% of cases, leading to both the highest number of true positives and the lowest number of false positive detections. Although simultaneously searching for transits while modeling correlated noise is expected to be computationally expensive, we make our algorithm tractable and available as the \textsf{JAX}-powered Python package \href{https://github.com/lgrcia/nuance}{\nuancecode{}}, allowing its use on distributed environments and GPU devices. Finally, we explore the prospects offered by the \nuancemethod{} formalism, and its use to advance our knowledge of planetary systems around active stars, both using space-based surveys and sparse ground-based observations. \href{https://github.com/lgrcia/nuance}{\color{linkcolor}\faCode}\,\,\href{https://github.com/lgrcia/paper-nuance}{\color{linkcolor}\faFileTextO}\,\,\href{https://nuance.readthedocs.io}{\color{linkcolor}\faBook}
\end{abstract}

@@

In this paper, we identify regions of light curves morphological parameter space for which a full-fledged transit search is necessary, and we present \nuancemethod{}\footnote{Throughout the paper, \nuancemethod{} written in italics refers to the algorithm, while \nuancecode{} in sans-serif refers to its implementation published as a Python package in its version 0.5.2 (see \autoref{package}).}, a method to search for transit signals while simultaneously modeling correlated noises in a tractable way. In \autoref{issues}, we describe the effect of correlated noise on transit light curves and the effect of its detrending on transit signals detectability. In \autoref{nuance}, we present \nuance{}, and the two main steps on which this method is based: the linear search and the periodic search. In \autoref{results}, we test the performance of \nuance{} on a wide variety of cases, and compare our method to commonly used transit search algorithms. This include transits injected in synthetic datasets, but also in the TESS light curves of 438 rapidly-rotating M dwarfs. Finally, in \autoref{discussion}, we discuss the results and the limitations of \nuance{}, before concluding in \autoref{conclusion}.\\\\
This paper was compiled from source code available in a git-versioned repository (\href{https://github.com/lgrcia/paper-nuance}{\color{linkcolor}\faFileTextO}) and used the \textsf{snakemake}\footlink{https://snakemake.readthedocs.io} workflow management tool. Source code of the \textsf{nuance} \textsf{Python} package is also available in a git-versioned repository (\href{https://github.com/lgrcia/nuance}{\color{linkcolor}\faCode}) together with an online documentation (\href{https://nuance.readthedocs.io}{\color{linkcolor}\faBook}). Finally, we provide links to the source code used to produce each figure (inidcated by the {\color{linkcolor}\codeicon} icon).

@@

        \includegraphics[width=\linewidth]{./principle_dataset_decomposed.pdf}
        \caption{The top plot shows the simulated flux time series of an active star drawn from the GP model described in \autoref{light_curves_simulations}. This dataset corresponds to an observation of 4 days with an exposure time of 2 minutes. The mean of this signal consists of a periodic transit signal of period $P=0.7$ days, duration $D=1.2$ hours and depth $\Delta=2\%$ (dark gray) plus an instrumental signal (blue). Correlated noise in the form of stellar variability is simulated with an SHO kernel of hyperparameters $\omega = \pi / 6 D$, $Q=45$ and $\sigma = \Delta$. Finally, white noise with a standard deviation of $0.1\%$ is added to the diagonal of the covariance matrix. \codelink{workflows/principle/principle.ipynb}}

@@

One way to quantify the detectability of a unique transit signal is to compute its signal-to-noise-ratio expressed as 
\begin{equation*}
  SNR= \frac{\Delta}{\sigma}\sqrt{n},
\end{equation*}
where $\Delta$ is the transit depth, $\sigma$ the measurements uncertainty (assuming homoscedasticity), and $n$ the number of points within transit. Although this metric is useful to assess the strength of the transit signal given a certain photometric precision, it does not account for the presence of correlated noise. However, instrumental and other astrophysical signals will necessarily affect the detectability of transits in realistic light curves \citep{Pont2006}. For this reason, the combined differential photometric precision (CDPP; \citealt{Jenkins2010}) metric was developed, and used in the context of the Kepler mission to assess the level of correlated noise in light curves, affecting the detectability of transits with a given duration. 
% Ultimately, this metric was essential to estimate the yield of the survey, and helped to achieve one of the primary goal of the mission: to measure the occurrence rate of Earth-size planets orbiting in the habitable zones of sun-like stars \citep{}.
The CDPP is computed by decomposing the data in the time-frequency domain using wavelets, and measures the significance of a distorted transit signal in the whitened data. For a given transit duration, the CDPP is then a measure of the noise remaining after filtering the light curve in each frequency band, taking into account the presence of nonstationary correlated noise of a given timescale (see \citealt{Jenkins2010} for more details). Hence, we can estimate the significance of the transit signal accounting for the presence of correlated noise as
\begin{equation}\label{eq:snr}
  SNR= \frac{\hat{\Delta}}{CDPP_D},
\end{equation}
where $CDPP_D$ is the CDPP computed for a given transit duration $D$ and $\hat{\Delta}$ is the transit depth after detrending. For simplicity, the CDPP is computed using the method from \cite{Gilliland2011} implemented in the \textsf{lightkurve} Python package \citep{Lightkurve} and $\hat{\Delta}$ is the depth obtained by solving \autoref{eq:linear_model} with the last column of $\bm{X}$ containing a normalized transit model.
\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=0.6\linewidth]{../workflows/plot_issues/figures/issue1.pdf}
        \caption{Illustration of the effect of correlated noise on a single transit SNR. An hour-long transit of depth 0.5\% is generated on top of white noise (standard deviation of 0.15\%) as part of a 24-hours observation with an exposure time of 1 minute (top). Then, in the bottom plot, correlated noise is added, generated using a GP with an SHO kernel of hyperparameters $\omega=60$, $Q=0.5$ and $\sigma=\Delta/4$ . The SNR on the right of each light curve is computed using \autoref{eq:snr}. Models used to simulate these data are provided in \autoref{light_curves_simulations}. \codelink{workflows/plot_issues/scripts/plot_issue_1.py}}
        \label{fig:issue1}
    \end{centering}
\end{figure}
\autoref{fig:issue1} shows the SNR from \autoref{eq:snr} computed for a unique transit observed in the absence (gray) and presence of correlated noise (red). As illustrated in this figure, the presence of correlated noise strongly decreases the transit signal SNR, which would ultimately limit its detectability.

@@

        \caption{In each plot on the left, correlated noises simulated using a GP with an SHO kernel are added (blue line) to an original light curve containing a transit of depth $\Delta = 0.8\%$, duration $D=0.1$ day, and white noise with a standard deviation of 0.15\% (top left). The hyperparameters of the kernel are fixed to $Q=10$ and $\sigma=0.8\%$, with the pulsation $\omega$ increasing from top to bottom. The resulting stellar variability signals are then reconstructed using (in purple) Tukey's bi-weight filter with an optimal window size of $3\times D$  (see \citealt{wotan}), and (in red) a GP with the same kernel used to simulate the data. In each case the variability is reconstructed, subtracted, and the transit SNR computed using \autoref{eq:snr}. \codelink{workflows/plot_issues/scripts/plot_issue_2.py}}

@@

        \caption{Transit SNR computed for 10 000 differential light curves with varying morphological characteristics. All light curves span 2.8 days with an exposure time of 2 minutes and contain the same transit signal of  duration $D=1$ hour, depth of 0.5\% and white noise with a standard deviation of 0.1\%. Light curves at the top and right side of the central plot are shown with their corresponding $\tau$ and $\delta$ values, which again correspond to the relative timescale of stellar variability against the transit duration, and the relative amplitude of stellar variability against the transit depth. The two central plots show the transit SNR after detrending the light curves with one of two methods. On the left, light curves are detrended using Tukey's bi-weight filter with an optimal window size of $3\times D$ (see \citealt{wotan}), while on the right, light curves are detrended using a GP with the same kernel used to simulate the data. \codelink{workflows/cleaning_snr/scripts/plot.py}}

@@

The goal of the linear search is to compute the likelihood $p(\bm{f} \vert T_0 , D, \Delta)$ for a grid of epochs $\set{T_i}_i$ and durations $\set{D_j}_j$. For each pair $(T_i, D_j)$ the transit depth $\Delta_{i, j}$ is linearly solved, which leads to the set of maximum likelihoods
\begin{equation*}
    \set{\ln\mathcal{L}_{i,j}}_{i, j} = \set{\ln p(\bm{f} \vert T_i ,D_j, \Delta_{i, j})}_{i, j}.
\end{equation*}
An example of such a grid of likelihoods is shown in \autoref{fig:linear_search}. To prepare for the next step, uncertainties on the depths $\sigma_{i,j}$ are also computed using \autoref{eq:ls} and stored.
\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_linear_search.pdf}
        \caption{Principle and output of the linear search. The simulated dataset (top) corresponds to the one shown and described in \autoref{fig:app_principle_dataset}. First, a set of durations and depths $\set{T_i, D_j}_{i,j}$ is generated. For each pair of indices $(i,j)$, the likelihood $\ln p(\bm{f} \vert T_i ,D_j, \Delta_{i,j})$ is computed using the parameters from \autoref{eq:ls} and the expression of \autoref{eq:linear_search_ll}. This process yields the grid of log-likelihoods $\ln\mathcal{L}$ (bottom plot), as well as the $\set{\Delta_{i,j}, \sigma_{i,j}}_{i, j}$ transit depths and errors inferred linearly using \autoref{eq:ls}. \codelink{workflows/principle/principle.ipynb}}
        \label{fig:linear_search}
    \end{centering}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_periodic_1.pdf}\\
    \caption{Periodic search likelihood $\mathcal{P}(P)$ computed for different trial periods $P$. Notice how the maximum value of $\mathcal{P}$ for the alias period $P=0.35$ day (left plot) is lower than for $P=2\times 0.35=0.7$ day, a result of combining the log-likelihoods using \autoref{eq:result} instead of \autoref{eq:attempt}, in favor of individual transits matching a common depth $\Delta$. \codelink{workflows/principle/principle.ipynb}}
    \label{fig:periodic_search}
\end{figure}

@@

Using \autoref{eq:result}, we can now compute $\ln\mathcal{P}$ for a range of periods and phases, and build a transit search periodogram using \autoref{eq:periodogram}. This has two disadvantages: First, each likelihood $p(\bm{f} \vert T_0, D, \Delta)$ estimated during the linear search is computed using N measurements. Hence, combining transits in the periodic search, through $\Delta_k$, $\sigma_k$ and the product of $K$ likelihoods $\set{\mathcal{L}_k}_k$ (see \autoref{eq:result}), artificially leads to a likelihood involving up to $N\times K$ measurements. For this reason, one has to normalize each likelihood $\ln\mathcal{P}$, by keeping track of the number of points used to compute each of them, which differs from one phase to another. Second, the maximum value of the likelihood $\ln\mathcal{P}$ is relative to a given dataset, so that a more intuitive and absolute metric must be used to relate to the transit signal detection (such as the Signal Detection Efficiency in \citealt{bls}). This motivates a final step to produce an interpretable transit search periodogram $\mathcal{Q}$. 
\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=\linewidth]{../workflows/principle/figures/principle_Q.pdf}
        \caption{For each period $P$, the joint likelihood $\mathcal{P}(P)$ is computed using \autoref{eq:result}, and the value of the maximum likelihood transit SNR retained as $\mathcal{Q}(P)$. \codelink{workflows/principle/principle.ipynb}}
        \label{fig:periodogram}
    \end{centering}
\end{figure}
For any period P, instead of taking $ \mathcal{Q}(P)$ as the maximum value of $\ln\mathcal{P}$, we compute the maximum likelihood parameters
\begin{equation}\label{eq:phi0}
    (\phi_0 ,D) = \argmax_{\phi_i, D_j} \set{\ln p(\bm{f} \vert P, \phi_i, D_j)}_{i, j}
\end{equation}
and define $\mathcal{Q}(P)$ as being equal to the SNR of the transit of period $P$, epoch $T_0 = \phi_0 \times P$, duration $D$ and depth $\Delta$, i.e.
\begin{equation*}
    \mathcal{Q}(P) = \frac{\Delta}{\sigma},
\end{equation*}
where $\Delta$ and $\sigma$ are obtained using \autoref{eq:result} with the last column of $X$ containing a periodic transit signal of period $P$, epoch $T_0$, duration $D$ and depth $1$. This process and the resulting periodogram $\mathcal{Q}$ are shown in \autoref{fig:periodogram}.\\\\
Hence, periodic transit of period $P$ with the maximum SNR, i.e. maximizing $\mathcal{Q}$, is adopted as the best candidate, basing the confidence in this signal through its SNR. The parameters of this transit are the period $P$, epoch $T_0 = \phi_0 P$, duration $D$ (\autoref{eq:phi0}), and depth $\Delta$ with error $\sigma$ (given by \autoref{eq:result}).\\\\
\subsection{An open-source python package}\label{package}
The methods presented in this paper are made available through the \nuancecode{} open-source Python package hosted at \href{https://github.com/lgrcia/nuance}{https://github.com/lgrcia/nuance}, released on the Python Package Index\footlink{https://pypi.org/project/nuance/} and with documentation and tutorials hosted at \href{https://nuance.readthedocs.io}{https://nuance.readthedocs.io}. All following mentions of \nuancecode{} refer to its \review{version 0.6.0 \citep{zenodo}.}{} \\\\
To instantiate a search, a user can start by creating a \textsf{Nuance} object with
\begin{lstlisting}[language=Python]
from nuance import Nuance

@@

\begin{figure}[H]
    \begin{centering}
        \includegraphics[width=\linewidth]{../workflows/control_test_bls/figures/control_test.pdf}
        \caption{Binned statistics of the injection-recovery of 4000 transit signals in a flat light curve with only white noise using \textsf{BLS} and \nuancecode{}. The color scale indicates the recovery of transits in the corresponding (period, depth) parameter space, white for a full recovery and black for no detection. \codelink{workflows/control_test_bls/scripts/plot.py}}
        \label{fig:control}
    \end{centering}
\end{figure}

@@

        \caption{Rate of true positives for the 4000 simulated light curves. The color scale represents the fraction of true positives, white if all injected transits are recovered in a given portion of the $(\tau, \delta)$ parameter space, black if none are recovered. \codelink{workflows/synthetic-injection-recovery/scripts/plot.py}}

@@

    \includegraphics[width=\linewidth]{cleaned_1019692.pdf}
    \caption{Trimmed and cleaned single-sector light curve of the target TIC 1019692. The top plot shows how much of the data is truncated and sigma clipped, resulting in a quasi-continuous light curve shown in the bottom plot. On this bottom plot, the black line corresponds to the mean of the GP model (with hyperparameters optimized here on the cleaned light curve). \codelink{workflows/tess_injection_recovery/scripts/plot_lc.py}}

@@

        \caption{Rate of true and false positives for all methods as a fraction of detectable transit signals. \codelink{workflows/tess_injection_recovery/scripts/plot_tpfp.py}}

@@

        \caption{Rate of true and false positives of \nuancecode{} compared to other methods, as a function of the relative parameters $\tau$ and $\delta$. The solid blue line corresponds to the maximum of true positives among all methods on the top panel (dominated \review{by the \texttt{harmonics+BLS} method}), and the minimum of false positives on the bottom panel (waned by the \texttt{GP+BLS} method, all other methods being above \nuancecode{}). \codelink{workflows/tess_injection_recovery/scripts/plot_tpfp.py}}

@@

        \caption{Rate of true and false positives binned over the full parameter space $(\tau, \delta)$ for all methods and all targets. \codelink{workflows/tess_injection_recovery/scripts/plot_tpfp.py}}

@@

        \caption{\review{Rate of true and false positives of \nuancecode{} compared to other methods, as a function of the injected planet's radius. The solid blue line corresponds to the maximum of true positives among all other methods than \nuance{} (dominated by the \texttt{harmonics+BLS} method).}{} \codelink{workflows/tess_injection_recovery/radii.ipynb}}
